---
title: "Lab 7 Solutions"

date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---
Name: Kaushik Raj V Nadar (208160499)

# Problem 1

P(1) Consider the Webster, Gunst, and Mason (WGM) data set given in Table 9.4 of Montgomery Book.

Load Data
```{r}
library(readxl)
dataset <- read_excel("Webster_Data.xlsx")

WGM_data <- as.data.frame(dataset)
```

(a) Normalise the data set by subtracting the associated mean and dividing by the square root of (n - 1) × the sample variance of each variables.

```{r}
normalize_data <- function(data) {
  n <- nrow(data)
  norm_data <- scale(data, center = TRUE, scale = sqrt((n - 1) * apply(data, 2, var)))
  return(norm_data)
}

# Assuming WGM_data is the dataset from Table 9.4 in Montgomery
normalized_WGM_data <- normalize_data(WGM_data)
normalized_WGM_data
```

(b) Find the VIFs
```{r}
calculate_vif_manual <- function(X) {
  # Compute the X^T X matrix
  XtX <- t(X) %*% X
  
  # Compute the inverse of (X^T X)
  XtX_inv <- solve(XtX)
  
  # The diagonal elements of the inverse of X^T X give the VIF
  vifs <- diag(XtX_inv)
  
  return(vifs)
}


X <- as.matrix(normalized_WGM_data[,-1])  # Exclude the response variable

vifs <- calculate_vif_manual(X)
vifs
```

(c) Find the condition number.
```{r}
XtX <- t(X) %*% X
eigen_values <- eigen(XtX)$values
condition_number <- max(eigen_values) / min(eigen_values)
condition_number
```

(d) Find the eigen vectors and verify if it matches with the eigen vectors given in Table 9.6 of Montgomery
Book.
```{r}
eigen_vectors <- eigen(XtX)$vectors
eigen_vectors
```

# Problem 2

P(2) Find variance Decomposition Proportions for the WGM data and verify if it matches with the results reported in Table 9.7 of Montgomery Book.
```{r}
calculate_vdp <- function(eigen_vectors, eigen_values, vifs) {
  p <- ncol(eigen_vectors)  # Number of predictors
  vdp_matrix <- matrix(0, nrow = p, ncol = p)  # Create a matrix to store VDPs
  
  # Loop over each predictor and each eigenvalue
  for (j in 1:p) {
    for (i in 1:p) {
      t_ji <- eigen_vectors[j, i]  # Element of the eigenvector matrix
      lam_i <- eigen_values[i]  # Square root of the eigenvalue
      vdp_matrix[i, j] <- (t_ji^2 / lam_i) / vifs[j]  # Calculate the VDP
    }
  }
  
  return(vdp_matrix)
}

# Compute the Variance Decomposition Proportions
vdp <- calculate_vdp(eigen_vectors, eigen_values, vifs)

# Display the VDP matrix
vdp
```