---
title: "Lab Assignment 11"
format: pdf
editor: visual
author: "Kaushik Raj V Nadar (208160499)"
date: today
---

# Problem 1

Consider the Aircraft Damage Data in which the response variable is the number of locations where damage was inflicted on the aircraft. The regressor x1 is an indicator variable (type of aircraft), and the other regressors x2 and x3 are bomb load (in tons) and total months of aircrew experience. Fit Poisson regression model and test the hypothesis for the significance of regressors.

#### Create the dataset

```{r}
aircraft_data <- data.frame(
  y = c(0,1,0,0,0,0,1,0,0,2,1,1,1,1,2,3,1,1,1,2,0,1,1,2,
        5,1,1,5,5,7),
  x1 = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,1,1,1,1,1,1,1,1,
         1,1,1,1,1,1),
  x2 = c(4,4,4,5,5,5,6,6,6,7,7,7,8,8,8,7,7,7,10,10,10,12,
         12,12,8,8,8,14,14,14),
  x3 = c(91.5,84.0,76.5,69.0,61.5,80.0,72.5,65.0,57.5,
         50.0,103.0,95.5,88.0,80.5,73.0,
         116.1,100.6,85.0,69.4,53.9,112.3,96.7,81.1,
         65.6,50.0,120.0,104.4,88.9,73.7,57.8)
)
```

#### Fit Poisson regression model

```{r}
poisson_model <- glm(y ~ x1 + x2 + x3, 
                     family = poisson(link = "log"), 
                     data = aircraft_data)
```

#### Display model summary

```{r}
summary(poisson_model)
```

#### Calculate pseudo R-squared (1 - residual deviance/null deviance)

```{r}
pseudo_r2 <- 1 - poisson_model$deviance/poisson_model$null.deviance
```

#### Perform analysis of deviance

```{r}
anova_result <- anova(poisson_model, test = "Chisq")
```

#### Print additional model statistics

```{r}
cat("\nModel Statistics:\n")
cat("Null Deviance:", poisson_model$null.deviance, 
    "on", poisson_model$df.null, "degrees of freedom\n")
cat("Residual Deviance:", poisson_model$deviance, 
    "on", poisson_model$df.residual, "degrees of freedom\n")
cat("AIC:", poisson_model$aic, "\n")
cat("Pseudo R-squared:", pseudo_r2, "\n")
```

#### Display ANOVA results

```{r}
print(anova_result)
```

#### Calculate and print odds ratios and confidence intervals

```{r}
coef_exp <- exp(coef(poisson_model))
conf_int <- exp(confint(poisson_model))
coef_table <- cbind(Estimate = coef_exp, 
                    CI_lower = conf_int[,1], 
                    CI_upper = conf_int[,2])

cat("\nExponentiated Coefficients (Rate Ratios) with 95% CI:\n")
print(coef_table)
```

#### Perform Walt's test for analysis of deviance

```{r}
library(lmtest)
walt_result <- waldtest(poisson_model)
cat("\nWald's Test Results:\n")
print(walt_result)

```

# Problem 2

Following Table contains data from an experiment conducted to investigate the three factors x1 = length, x2 = amplitude, and x3 = load on the cycles to failure y of worsted yarn. The regressor variables are coded, and readers who have familiarity with designed experiments will recognize that the experimenters here used a 33 factorial design. Fit a GLM using the gamma distribution and the log link.

```{r}
# Create the dataset
yarn_data <- data.frame(
  y = c(674, 1414, 3636, 338, 1022, 1568, 170, 442, 1140, 370, 1198, 
        3184, 266, 620, 1070, 118, 332, 884, 292, 634, 2000, 210,
        438, 566, 90, 220, 360),
  x1 = c(-1, 0, 1, -1, 0, 1, -1, 0, 1, -1, 0, 1, -1, 0, 1, -1, 0, 1,
         -1, 0, 1, -1, 0, 1, -1, 0, 1),
  x2 = c(-1, -1, -1, 0, 0, 0, 1, 1, 1, -1, -1, -1, 0, 0, 0, 1, 1, 1,
         -1, -1, -1, 0, 0, 0, 1, 1, 1),
  x3 = c(-1, -1, -1, -1, -1, -1, -1, -1, -1, 0, 0, 0, 0, 0, 0, 0, 0, 
         0, 1, 1, 1, 1, 1, 1, 1, 1, 1)
)

# Fit Gamma regression model with log link
gamma_model <- glm(y ~ x1 + x2 + x3, family = Gamma(link = "log"), data = yarn_data)

# Display model summary
summary(gamma_model)

# Calculate pseudo R-squared (1 - residual deviance/null deviance)
pseudo_r2 <- 1 - gamma_model$deviance/gamma_model$null.deviance

# Perform analysis of deviance
anova_result <- anova(gamma_model, test = "Chisq")

# Print additional model statistics
cat("\nModel Statistics:\n")
cat("Null Deviance:", gamma_model$null.deviance, 
    "on", gamma_model$df.null, "degrees of freedom\n")
cat("Residual Deviance:", gamma_model$deviance, 
    "on", gamma_model$df.residual, "degrees of freedom\n")
cat("AIC:", gamma_model$aic, "\n")
cat("Pseudo R-squared:", pseudo_r2, "\n")

# Display ANOVA results
cat("\nAnalysis of Deviance Table:\n")
print(anova_result)

# Calculate and print odds ratios and confidence intervals
coef_exp <- exp(coef(gamma_model))
conf_int <- exp(confint(gamma_model))
coef_table <- cbind(Estimate = coef_exp, 
                    CI_lower = conf_int[,1], 
                    CI_upper = conf_int[,2])

cat("\nExponentiated Coefficients (Rate Ratios) with 95% CI:\n")
print(coef_table)

walt_result <- waldtest(poisson_model)
cat("\nWald's test results:\n")
print(walt_result)
```
